---
title: 'Broman Jax Figures'
author: "Brian S. Yandell"
date: "1/3/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This just reproduces figures from Karl Broman's Jax talk. Here is a list of figures:

- DH_bw.jpg
- backcross_bw.jpg
- genoprob1_bw.jpg
- genoprob2_bw.jpg
- genoprob3_bw.jpg
- genoprob4_bw.jpg
- genoprob5_bw.jpg
- genoprob6_bw.jpg
- intercross_bw.jpg
- mixtures_bw.jpg
- structure_bw.jpg
- loddist_bw.jpg
- permtest_bw.jpg
- perm_hist_bw.jpg
- lodsuppint_bw.jpg

```{r}
library(qtlbook)
source("func.R")
source("colors_bw.R")
bropath <- "~/Documents/Talks/broman_karl/Talk_JAXsysgen"
bw <- TRUE
```

Trick to create jpeg instead of pdf:

```{r}
pdf <- function(file, width, height, onefile, ...) {
  jpeg(file = stringr::str_replace(file, ".pdf", ".jpg"), width = width * 96, height = height * 96, ...)
}
```

## Design Talk

```{r}
source(file.path(bropath, "Intro/R", "backcross_fig.R"))
source(file.path(bropath, "Intro/R", "intercross_fig.R"))
source("DH_fig.R")
```

## Models Talk

```{r}
source(file.path(bropath, "Intro/R", "structure_fig.R"))
source(file.path(bropath, "Intro/R", "genoprob_fig.R"))
source(file.path(bropath, "Intro/R", "mixtures_fig.R"))
```

```{r}
if(!dir.exists("Rcache")) {
  dir.create("Rcache")
}
source(file.path(bropath, "Intro/R", "loddist_fig.R"))
source(file.path(bropath, "Intro/R", "permtest.R"))
```

The `perm_hist.R` seems to get stuck. Had to do by hand.

```{r}
source(file.path(bropath, "Intro/R", "alod_fig.R"))
source(file.path(bropath, "Intro/R", "perm_hist.R"))
```

```{r}
source(file.path(bropath, "Intro/R", "lodsupint_fig.R"))
```

